\begin{abstract}
% Compulsory
% 1 page max. Tony: "third to half a page ideal probably"
% Anandha: "Don't put a photo of your group in the executive summary or
% something like that!"
% Wael: I would say it might be appropriate to do something like that in the
% presentation though. We can discuss and see.

Attractor Neural networks have been a subject of intensive research in the past 30 years. They have been proposed as a model of associative memory and have been used in ranging from facial and speech recognition to modelling biological activities of the human brain.


The aim of our project was to attain a deeper understanding of these networks by running various simulations and investigating their properties. For this purpose, we have demonstrated their importance by implementing an image recognition software. Furthermore, we have also used them as an aid for modelling a psychological concept known as Attachment Theory, in order to better understand the human mind and how to cure various mental disorders.


Our project attempts to model, at a metaphorical level, an individual's emotional mindset, and how it is formed due to the influences of experiences, which manifest as memories, encountered in early life. We have implemented several mathematical models of neural networks, that can simulate various functions of the brain. Amongst these, two of them are of uttermost importance: learning and recalling information.


We have used our networks to analyse and quantify the influence of memories, analysing factors affecting their probability of being recalled, or how they can be either forgotten or reinforced. The latter is extremely important for the psychological studies, since it paves the way for curing some mental disorders whose roots lie in negative experiences encountered early in life.

% should we include some of the results as well?
\end{abstract}
